---
title: "Hands-on Exercise 4"
author: "SMLurker"
---

# Visual Statistical Analysis

## Installing and launching R packages

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## Import Data

```{r}
exam <- read.csv("data/Exam_data.csv")
```

## One-sample test: gghistostats() method

#### In the code chunk below, gghistostats() is used to to build an visual of one-sample test on English scores.

#### Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

## Two-sample mean test: ggbetweenstats()

#### In the code chunk below, ggbetweenstats() is used to build a visual for two-sample mean test of Maths scores by gender.

#### Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

## Oneway ANOVA Test: ggbetweenstats() method

#### In the code chunk below, ggbetweenstats() is used to build a visual for One-way ANOVA test on English score by race.

#####  "ns" → only non-significant

#####  "s" → only significant

#####  "all" → everything

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## Significant Test of Correlation: ggscatterstats()

#### In the code chunk below, ggscatterstats() is used to build a visual for Significant Test of Correlation between Maths scores and English scores.

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

## Significant Test of Association (Depedence) : ggbarstats() methods

#### In the code chunk below, the Maths scores is binned into a 4-class variable by using cut().

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
```

#### In this code chunk below ggbarstats() is used to build a visual for Significant Test of Association

```{r}
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

## Installing and loading the required libraries for Visualising Models

```{r}
pacman::p_load(readxl, performance, parameters, see)
```

## Importing Excel file: readxl methods

#### In the code chunk below, read_xls() of readxl package is used to import the data worksheet of ToyotaCorolla.xls workbook into R. Notice that the output object car_resale is a tibble data frame.

```{r}
car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       "data")
car_resale
```

## Multiple Regression Model using lm()

#### The code chunk below is used to calibrate a multiple linear regression model by using lm() of Base Stats of R.

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
model
```

## Model Diagnostic: checking for multicolinearity

#### In the code chunk, check_collinearity() of performance package.

```{r}
check_collinearity(model)
```

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

## Model Diagnostic: checking normality assumption

#### In the code chunk, check_normality() of performance package.

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
```

```{r}
check_n <- check_normality(model1)
```

```{r}
plot(check_n)
```

## Model Diagnostic: Check model for homogeneity of variances

#### In the code chunk, check_heteroscedasticity() of performance package.

```{r}
check_h <- check_heteroscedasticity(model1)
```

```{r}
plot(check_h)
```

## Model Diagnostic: Complete check

#### We can also perform the complete by using check_model().

```{r}
check_model(model1)
```

## Visualising Regression Parameters: see methods

#### In the code below, plot() of see package and parameters() of parameters package is used to visualise the parameters of a regression model.

```{r}
plot(parameters(model1))
```

## Visualising Regression Parameters: ggcoefstats() methods

#### In the code below, ggcoefstats() of ggstatsplot package to visualise the parameters of a regression model.

```{r}
ggcoefstats(model1, 
            output = "plot")
```

# Visualising Uncertainty

## Visualizing the uncertainty of point estimates

```{r}
pacman::p_load(tidyverse, plotly, crosstalk, DT, ggdist, gganimate)
```

```{r}
exam <- read_csv("data/Exam_data.csv")
```

## Visualizing the uncertainty of point estimates: ggplot2 methods

#### The code chunk below performs the followings:

#####  group the observation by RACE,

#####  computes the count of observations, mean, standard deviation and standard error of Maths by RACE, and

#####  save the output as a tibble data table called my_sum.

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))
```

#### Note: For the mathematical explanation, please refer to Slide 20 of Lesson 4.

#### Next, the code chunk below will

```{r}
knitr::kable(head(my_sum), format = 'html')
```

## Visualizing the uncertainty of point estimates: ggplot2 methods

#### The code chunk below is used to reveal the standard error of mean maths score by race.

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean 
          maths score by rac")
```

## Visualizing the uncertainty of point estimates: ggdist methods

#### In the code chunk below, stat_pointinterval() of ggdist is used to build a visual for displaying distribution of maths scores by race.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +   #<<
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

## Visualizing the uncertainty of point estimates: ggdist methods

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval(
    show.legend = FALSE) +   
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

## Visualizing the uncertainty of point estimates: ggdist methods

#### In the code chunk below, stat_gradientinterval() of ggdist is used to build a visual for displaying distribution of maths scores by race.

```{r}
#| warning: false
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")
```

## Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)

#### Step 1: Installing ungeviz package

```{r}
devtools::install_github("wilkelab/ungeviz")
```

#### Note: You only need to perform this step once.

#### Step 2: Launch the application in R

```{r}
library(ungeviz)
```

```{r}
ggplot(data = exam, 
       (aes(x = factor(RACE), y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, width = 0.05), 
    size = 0.4, color = "#0072B2", alpha = 1/2) +
  geom_hpline(data = sampler(25, group = RACE), height = 0.6, color = "#D55E00") +
  theme_bw() + 
  # `.draw` is a generated column indicating the sample draw
  transition_states(.draw, 1, 3)
```

## Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)

```{r}
ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)
```

# Funnel Plots for Fair Comparisons

## Installing and Launching R Packages

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

## Importing Data

#### In this section, COVID-19_DKI_Jakarta will be used. The data was downloaded from Open Data Covid-19 Provinsi DKI Jakarta portal. For this hands-on exercise, we are going to compare the cumulative COVID-19 cases and death by sub-district (i.e. kelurahan) as at 31st July 2021, DKI Jakarta

```{r}
covid19 <- read_csv("data/COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)
```

## FunnelPlotR methods: The basic plot

```{r}
funnel_plot(
  numerator = covid19$Positive,
  denominator = covid19$Death,
  group = covid19$`Sub-district`
)
```

#### Things to learn from the code chunk above.

##### \ group in this function is different from the scatterplot. Here, it defines the level of the points to be plotted i.e. Sub-district, District or City. If Cityc is chosen, there are only six data points.
##### \ By default, data_typeargument is “SR”.
##### \ limit: Plot limits, accepted values are: 95 or 99, corresponding to 95% or 99.8% quantiles of the distribution.

## FunnelPlotR methods: Makeover 1

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",     #<<
  xrange = c(0, 6500),  #<<
  yrange = c(0, 0.05)   #<<
)
```

#### Things to learn from the code chunk above. + data_type argument is used to change from default “SR” to “PR” (i.e. proportions). + xrange and yrange are used to set the range of x-axis and y-axis

## FunnelPlotR methods: Makeover 2

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```

#### Things to learn from the code chunk above.

##### \ label = NA argument is to removed the default label outliers feature.
##### \ title argument is used to add plot title.
##### \ x_label and y_label arguments are used to add/edit x-axis and y-axis titles.

## Computing the basic derived fields

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
```

```{r}
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)
```

## Calculate lower and upper limits for 95% and 99.9% CI

```{r}
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

## Plotting a static funnel plot

```{r}
p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p
```

## Interactive Funnel Plot: plotly + ggplot2

```{r}
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly

```

