---
title: "Take-home Exercise 1"
author: "Wang Shengming"
date: "06 May 2023"
data_modified: "`r Sys.Date()`"
execute:
  warning: false
  echo: true
  eval: true
---

# 1. Setting the Scene

#### City of Engagement, with a total population of 50,000, is a small city located at Country of Nowhere. The city serves as a service centre of an agriculture region surrounding the city. The main agriculture of the region is fruit farms and vineyards. The local council of the city is in the process of preparing the Local Plan 2023. A sample survey of 1000 representative residents had been conducted to collect data related to their household demographic and spending patterns, among other things. The city aims to use the data to assist with their major community revitalization efforts, including how to allocate a very large city renewal grant they have recently received.

# 2. The Task

#### In this take-home exercise, you are required to apply the concepts and methods you had learned in Lesson 1-4 to reveal the demographic and financial characteristics of the city of Engagement, using appropriate static and interactive statistical graphics methods. This exercise requires a user-friendly and interactive solution that helps city managers and planners to explore the complex data in an engaging way and reveal hidden patterns. The data should be processed by using appropriate tidyverse family of packages and the statistical graphics must be prepared using ggplot2 and its extensions.

# 3. Data Description

#### For the purpose of this study, two data sets are provided. They are:

#### Participants.csv

##### Contains information about the residents of City of Engagement that have agreed to participate in this study.

######   participantId (integer): unique ID assigned to each participant.

######   householdSize (integer): the number of people in the participant's household

######   haveKids (boolean): whether there are children living in the participant's household.

######   age (integer): participant's age in years at the start of the study.

######   educationLevel (string factor): the participant's education level, one of: {"Low", "HighSchoolOrCollege", "Bachelors", "Graduate"}

######   interestGroup (char): a char representing the participant's stated primary interest group, one of {"A", "B", "C", "D", "E", "F", "G", "H", "I", "J"}. Note: specific topics of interest have been redacted to avoid bias.

######   joviality (float): a value ranging from \[0,1\] indicating the participant's overall happiness level at the start of the study.

#### FinancialJournal.csv

##### Contains information about financial transactions.

######   participantId (integer): unique ID corresponding to the participant affected

######   timestamp (datetime): the time when the check-in was logged

######   category (string factor): a string describing the expense category, one of {"Education", "Food", "Recreation", "RentAdjustment", "Shelter", "Wage"}

######   amount (double): the amount of the transaction

## 4. Install, launch R packages, load data and build data table

#### Reading the paticipants csv and build interactive paticipants table

```{r}
#| code-fold: true
#| code-summary: "Show the code"
pacman::p_load(tidyverse,dplyr,DT,plotly, ggdist,gganimate,ggstatsplot)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
p_data <- read_csv("data/Participants.csv") %>% 
  mutate_if(is.character, as.factor) %>% 
  distinct()
DT::datatable(p_data, class= "compact")
```

#### Modify the time style of the financial data, pivot the catogory and regroup, calculate the Savings, and build interactive FinancialJournal table

```{r}
#| code-fold: true
#| code-summary: "Show the code"
fin_data <- read_csv("data/FinancialJournal.csv") %>% 
  mutate_if(is.character, as.factor) %>% distinct() 
fin_data$timestamp <- format(as.Date(fin_data$timestamp), "%Y-%m")
f_data <- fin_data %>% group_by(participantId,timestamp,category) %>% 
  summarise(total_amount=sum(amount),
            .groups = 'drop') %>%
  as.data.frame()
f_data <- f_data %>%
  pivot_wider(.,names_from = "category", values_from = "total_amount") 
f_data <- replace(f_data, is.na(f_data), 0) %>%
    mutate(Saving = select(., Education:RentAdjustment) %>% rowSums(na.rm = TRUE))
DT::datatable(f_data, class= "compact")
```

## 5. Analysis and visualization

#### 5.1 Figure out the relationship of educationLevel and Joviality

###### Overall, we see from below that with the increasing of education level, people tend to have higher level of joviality as the lower bound and mean value of joviality has a trend to increase. Although we can also observe that for low level of education, the range is bigger that all other level, this might indicate longer social experience can still affect the upper bound of joviality, but enhancing education is still a more direct way to increase joviality level.

::: panel-tabset
## The plot

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
df <- f_data %>% group_by(participantId) %>% 
  summarise(total_edu=sum(Education),
            total_food=sum(Food),
            total_rec=sum(Recreation),
            total_shl=sum(Shelter),
            total_w=sum(Wage),
            total_rent=sum(RentAdjustment),
            total_saving=sum(Saving),
            .groups = 'drop') %>%
  as.data.frame()
df2 <- merge(x= df,y= p_data, by = "participantId")

my_sum <- df2 %>%
  group_by(educationLevel) %>%
  summarise(
    n=n(),
    mean=mean(joviality),
    sd=sd(joviality)
    ) %>%
  mutate(se=sd/sqrt(n-1))

ggplot(my_sum) +
  geom_errorbar(
    aes(x=educationLevel, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=educationLevel, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  scale_x_discrete(limits = c("Low", "HighSchoolOrCollege", "Bachelors", "Graduate")) +
  ggtitle("Standard error of mean 
          Joviality score by educationLevel")
```

## The code

```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
df <- f_data %>% group_by(participantId) %>% 
  summarise(total_edu=sum(Education),
            total_food=sum(Food),
            total_rec=sum(Recreation),
            total_shl=sum(Shelter),
            total_w=sum(Wage),
            total_rent=sum(RentAdjustment),
            total_saving=sum(Saving),
            .groups = 'drop') %>%
  as.data.frame()
df2 <- merge(x= df,y= p_data, by = "participantId")

my_sum <- df2 %>%
  group_by(educationLevel) %>%
  summarise(
    n=n(),
    mean=mean(joviality),
    sd=sd(joviality)
    ) %>%
  mutate(se=sd/sqrt(n-1))

ggplot(my_sum) +
  geom_errorbar(
    aes(x=educationLevel, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=educationLevel, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  scale_x_discrete(limits = c("Low", "HighSchoolOrCollege", "Bachelors", "Graduate")) +
  ggtitle("Standard error of mean 
          Joviality score by educationLevel")
```

## The statistics table

```{r}
summ <- my_sum[order(factor(my_sum$educationLevel, levels = c("Low", "HighSchoolOrCollege", "Bachelors", "Graduate"))),c("educationLevel","n","mean","sd")]
knitr::kable(head(summ), format = 'html')
```
:::

## 5.2 Figure out the relationship of householdSize and Joviality

#### From the plotting below, we see that number of householdSize does not have huge effect on Joviality but there is still some slight impact that household with 2 or more people tend to have a bit higher mean joviality than 1 person and 2 people have the hight mean.

::: panel-tabset

## The plot

```{r}
#| warning: false
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggbetweenstats(
  data = df2,
  x = householdSize, 
  y = joviality,
  type = "np",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
## The code

```{r}
#| warning: false
#| eval: false
#| fig-width: 4
#| fig-height: 4
ggbetweenstats(
  data = df2,
  x = householdSize, 
  y = joviality,
  type = "np",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
:::

## Distribution of savings

#### from the graph below, we can find that the saving of most people are below 50000 within the study period.

```{r}
gghistostats(
  data = df2,
  x = total_saving,
  type = "bayes",
  test.value = 60,
  xlab = "total_saving"
)
```

## And by plotting the ralationship of savings, interestGroups and joviality, we can see that no matter how much people save, they are still willing to join any kinds of interest groups. 

```{r}
plot_ly(data = df2, 
        x = ~total_saving, 
        y = ~joviality, 
        color = ~interestGroup)

```


## Conclusion

#### Therefore, in order to assist with the major community revitalization efforts, the very large city renewal grant they have recently received should be allocated as below:

###### \ Spend money on schools and universities to encourage people to gain knowledge and improve education qualification

###### \ Make dating apps or other method to help single people conveniently meet and connect with like-minded people

###### \ Hold activitied and interest groups to provide people with opportunities to meet people of same interests.